Spin canonical rings of log stacky curves

### 2015-07-07 curves-write-up/spin_cring.tex
surj arrow to normal arrow in g0 ex
minor example edits (e.g. n->d)

Next, we construct generators and relations for $R_L$ using those
of $R_{L_X}$. Note that we have a natural inclusion $\iota: R_{L_X}
\hookrightarrow R_L$ induced by the inclusions $H^0(\sx, L_X)
\hookrightarrow H^0(\sx, L).$ By Riemann Roch, we see there is some
element $y_{1, 3} \in (R_L)_3$ with $\ord_{P_1}(y_{1, 3}) = -1,$ not
in the image of the inclusion $\iota$. We claim that there exist $a_
1, a_2 \in k$ and a degree 4 polynomial $f(x_1, x_2) \in k[x_1, x_2]$
such that
\begin{align*}
	R_L \cong k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))
\end{align*}

First, note that $x_1, x_2, y_{1,3}$ generate all of $R_L$ from the
Generalized Max Noether Theorem from Voight and Zureick-Brown
\cite[Lemma 3.1.3]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx, 3L) \otimes H^0(\sx, (n - 3)L) \rightarrow H^0(\sx, nL)
\end{align*}

\noindent
are surjective for $n \geq 4$. A relation of the form $a_1 x_1 y_{1,
3} +a_2 x_2 y_{1, 3} + f(x_1, x_2) = 0$ must exist because $h^0(\sx
, 4L) - h^0(\sx, 4L_X) = 1,$ but $x_1 y_{1, 3}$ and $x_2 y_{1, 3}$
define two linearly independent elements with nontrivial image in
the 1-dimensional vector space $H^0(\sx, 4L) / H^0(\sx, 4L_X).$ So,
we obtain a surjection
\begin{align}
\label{align:surjection}
	k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)) 
	\rightarrow R_L.
\end{align}

To complete the example, it suffices to show there are no
additional relations. First, note that $a_1 x_1 y_{1, 3} + a_2 x_2
y_{ 1, 3} + f(x_1, x_2)$ is irreducible because there are no
relations among $x_1, x_2$ and $y_{1, 3}$ in lower degrees. Hence,
$k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} +
f(x_1, x_2))$ is integral and is 2-dimensional. Thus, the map
\eqref{align:surjection} defines a surjection from an integral
2-dimensional ring to a 2-dimensional ring. Therefore, it is an 
isomorphism.
\end{example}

@@

\begin{align*}
	&H^0 (\sx', \lfloor 3 \halfcan' \rfloor) \otimes H^0 (\sx', \lfloor
	(d - 3) \halfcan' \rfloor) \twoheadrightarrow H^0 (\sx', \lfloor
	d \halfcan' \rfloor) \text{ if } d \geq 12 \text{ and } 3 \nmid d \\
	&H^0 (\sx', \lfloor 7 \halfcan' \rfloor) \otimes H^0 (\sx', \lfloor
	(d - 7) \halfcan' \rfloor) \twoheadrightarrow H^0 (\sx', \lfloor
	d \halfcan' \rfloor) \text{ if } d \geq 6, d \not\equiv 0, 5
	\bmod 7
\end{align*}

>>>>

Next, we construct generators and relations for $R_L$ using those
of $R_{L_X}$. Note that we have a natural inclusion $\iota: R_{L_X}
\hookrightarrow R_L$ induced by the inclusions $H^0(\sx, dL_X)
\hookrightarrow H^0(\sx, dL)$ for each $d \geq 0$. By Riemann Roch,
we see there is some element $y_{1, 3} \in (R_L)_3$ with $\ord_{P_1}
(y_{1, 3}) = -1$, not in the image of the inclusion $\iota$. We
claim that there exist $a_1, a_2 \in k$ and a degree $4$ polynomial
$f(x_1, x_2) \in k[x_1, x_2]$ such that
\begin{align*}
	R_L \cong k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3}
	+ f(x_1,x_2))
\end{align*}

First, note that $x_1, x_2, y_{1, 3}$ generate all of $R_L$ from the
Generalized Max Noether Theorem for genus zero curves from Voight
and Zureick-Brown \cite[Lemma 3.1.1]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx, 3L) \otimes H^0(\sx, (d - 3)L) \rightarrow H^0(\sx, dL)
\end{align*}

\noindent
are surjective for $d \geq 4$. A relation of the form $a_1 x_1 y_{1,
3} + a_2 x_2 y_{1, 3} + f(x_1, x_2) = 0$ must exist because $h^0(\sx,
4L) - h^0(\sx, 4L_X) = 1,$ but $x_1 y_{1, 3}$ and $x_2 y_{1, 3}$
define two linearly independent elements with nontrivial image in
the 1-dimensional vector space $H^0(\sx, 4L) / H^0(\sx, 4L_X).$ So,
we obtain a surjection
\begin{align}
\label{align:surjection}
	k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} + f(x_1, x_2)) 
	\rightarrow R_L.
\end{align}

To complete the example, it suffices to show there are no
additional relations. First, note that $a_1 x_1 y_{1, 3} + a_2 x_2
y_{1, 3} + f(x_1, x_2)$ is irreducible because there are no
relations among $x_1, x_2$ and $y_{1, 3}$ in lower degrees. Hence,
$k[x_1, x_2, y_{1, 3}] / (a_1 x_1 y_{1, 3} + a_2 x_2 y_{1, 3} +
f(x_1, x_2))$ is integral and is 2-dimensional. Thus, the map
\eqref{align:surjection} defines a surjection from an integral
2-dimensional ring to a 2-dimensional ring. Therefore, it is an 
isomorphism.
\end{example}

@@

\begin{align*}
	&H^0 (\sx', \lfloor 3 \halfcan' \rfloor) \otimes H^0 (\sx', \lfloor
	(d - 3) \halfcan' \rfloor) \rightarrow H^0 (\sx', \lfloor
	d \halfcan' \rfloor) \text{ if } d \geq 12 \text{ and } 3 \nmid d \\
	&H^0 (\sx', \lfloor 7 \halfcan' \rfloor) \otimes H^0 (\sx', \lfloor
	(d - 7) \halfcan' \rfloor) \rightarrow H^0 (\sx', \lfloor
	d \halfcan' \rfloor) \text{ if } d \geq 6, d \not\equiv 0, 5
	\bmod 7
\end{align*}
